---
title: "Covid Casis Analysis Miami-Dade using C3.ai datalake"
output: html_notebook

---
title: "C3.ai COVID-19 Data Lake Quickstart in R"
output: html_notebook
---

# C3.ai COVID-19 Data Lake Quickstart in R  

Version 4.0 (June 23, 2020)

This R notebook shows some examples of how to access and use each of the [C3.ai COVID-19 Data Lake](https://c3.ai/covid/) APIs. These examples show only a small piece of what you can do with the C3.ai COVID-19 Data Lake, but will get you started with performing your own exploration. See the [API documentation](https://c3.ai/covid-19-api-documentation/) for more details.

Please contribute your questions, answers and insights on [Stack Overflow](https://www.stackoverflow.com). Tag `c3ai-datalake` so that others can view and help build on your contributions. For support, please send email to: [covid@c3.ai](mailto:covid@c3.ai).

To view an outline of this notebook, use the RStudio keyboard shortcut Control + Shift + O on Windows or Command + Shift + O on Mac. 

Import the tidyverse (version >= 1.3.0), httr, jsonlite, and fitdistrplus libraries to use this notebook.
```{r}
if (!require(tidyverse)) install.packages('tidyverse')
if (!require(httr)) install.packages('httr')
if (!require(jsonlite)) install.packages('jsonlite')
if (!require(fitdistrplus)) install.packages('fitdistrplus')
if (!require(fitdistrplus))install.packages("viridis")  # Install
library("viridis")           # Load

library(tidyverse)
library(httr)
library(jsonlite)
library(fitdistrplus)
library(dplyr)
library(readr) #To read the dataset
library(tidyverse) #For data manipulation and visual analytics
if (!require(correlationfunnel)) install.packages("correlationfunnel")
library(correlationfunnel)
library(DataExplorer)
if (!require(WVPlots)) install.packages("WVPlots")
library(WVPlots)
if (!require(ggthemes)) install.packages("ggthemes")
library(ggthemes)
if (!require(corrplot)) install.packages("corrplot")
library(corrplot)
if (!require(ROCR)) install.packages("ROCR")

library(ROCR)

library(caret)
if (!require(cowplot)) install.packages('cowplot')
if (!require(egg)) install.packages('egg')
if (!require(ggpubr)) install.packages('ggpubr')

library(cowplot)
library(egg)
library(ggpubr)
install.packages("rlist")
library(rlist)
```

## Helper methods for accessing the API

The helper methods in `c3aidatalake.R` convert a JSON response from the C3.ai APIs to a Tibble. You may wish to view the code in `c3aidatalake.R` before using the quickstart examples.
```{r}
source("c3aidatalake.R")
```

## Access OutbreakLocation data

OutbreakLocation stores location data such as countries, provinces, cities, where COVID-19 outbeaks are recorded. See the [API documentation](https://c3.ai/covid-19-api-documentation/#tag/OutbreakLocation) for more details and for a list of available locations.

```{r}
# Fetch facts about UnitedSates
locations <- fetch(
    "outbreaklocation",
    list(
      spec = list(
        filter = "id == 'UnitedStates'"
      )
    )
)

locations

locations %>% 
  dplyr::select(-location) %>% 
  unnest_wider(12086, names_sep = ".")
```





### Case counts

A variety of sources provide counts of cases, deaths, recoveries, and other statistics for counties, provinces, and countries worldwide.
```{r}
# Total number of confirmed cases, deaths, and recoveries in Miami-Dade County, Florida
today <- Sys.Date()

#cases for march
casecounts <- evalmetrics(
  "outbreaklocation",
  list(
    spec = list(
      ids = list("Miami-Dade_Florida_UnitedStates"),
      expressions = list("JHU_ConfirmedCases", "JHU_ConfirmedDeaths", "JHU_ConfirmedRecoveries"),
      start = "2020-03-01",
      end = "2020-03-30",
      #end = today,
      interval = "DAY"
    )
  )
)

#cases until system date
casecounts2 <- evalmetrics(
  "outbreaklocation",
  list(
    spec = list(
      ids = list("Miami-Dade_Florida_UnitedStates"),
      expressions = list("JHU_ConfirmedCases", "JHU_ConfirmedDeaths", "JHU_ConfirmedRecoveries"),
      start = "2020-03-01",
      #end = "2020-03-30",
      end = today,
      interval = "DAY"
    )
  )
)
```



Checking all data
```{r}
check_data <- function(casecountsx){
  
  casecounts%>%
    arrange(desc(dates))
  glimpse(casecounts)
  casecounts%>%
    plot_missing()
  
  casecountsfirst <-casecounts%>%
    filter(data>0)%>%
    filter(value_id == "JHU_ConfirmedCases" )
  casecountsfirst$dates[1]  
}
check_data(casecounts)
```

Plot these counts.
this plot is for Miami Dade with casecounts for the month of march
```{r}
plot1 <- casecounts %>% 
  ggplot(aes(dates, data, color = value_id))+
  geom_line() +
  scale_y_log10()+
  facet_wrap(vars(name)) +
  
 #theme_dark()
  #scale_color_viridis(discrete = FALSE, option = "C")+
  labs(
    x = "Date",
    y = "Count",
    color = "Metric"
  )
plot2 <- casecounts %>% 
  ggplot(aes(dates, data, color = value_id), palette= "jco") +
  geom_line() +
  facet_wrap(vars(name)) +
  labs(
    x = "Date",
    y = "Count",
    color = "Metric"
  )

#cowplot::plot_grid(plot1, plot2, align = "v", ncol = 2,nrow1 = 1)# rel_heights = c(0.25, 0.75))

figure <-ggarrange(plot1, plot2,ncol = 1,nrow = 2)
               # label.y = 1,
               # hjust = -0.5,
               # vjust = 1.5,
               # font.label = list(size = 14, color = "black", face = "bold", family = NULL),
               # widths = 1,
               # heights = 1)# heights = c(0.25, 0.75))
annotate_figure(figure,
                top = text_grob("Miami Dade First Case in March", color = "blue", face = "bold", size = 14),
                bottom = text_grob("Data source: \ C3.ai Covid Data Lake", color = "blue",
                                   hjust = 1, x = 1, face = "italic", size = 10),
                left = text_grob("Data in Linear and Logarithmic Scales", color = "blue", rot = 90),
                # right = "I'm done, thanks :-)!",
                fig.lab = "Figure 1", fig.lab.face = "bold"
                )
```

This function will plot casecounts
```{r}
plot_cases<-function (casecountsx) {
  #### total as of today
  plot1 <- casecountsx %>% 
    ggplot(aes(dates, data/1000, color = value_id))+
    geom_line() +
    scale_y_log10(labels = waiver(), breaks= c(10,100,1000,10000,100000,1000000,10000000))+
    facet_wrap(vars(name)) +
    
   #theme_dark()
    #scale_color_viridis(discrete = FALSE, option = "C")+
    labs(
      x = "Date",
      y = "Count (Thousands) - Log",
      color = "Metric"
    )
  plot2 <- casecountsx %>% 
    ggplot(aes(dates, data, color = value_id), palette= "jco") +
    geom_line() +
    facet_wrap(vars(name)) +
    labs(
      x = "Date",
      y = "Count - Linear",
      color = "Metric"
    )
  if (!require(cowplot)) install.packages('cowplot')
  if (!require(egg)) install.packages('egg')
  if (!require(ggpubr)) install.packages('ggpubr')
  
  library(cowplot)
  library(egg)
  library(ggpubr)
  #cowplot::plot_grid(plot1, plot2, align = "v", ncol = 2,nrow1 = 1)# rel_heights = c(0.25, 0.75))
  
  figure <-ggarrange(plot1, plot2,ncol = 1,nrow = 2)
                 # label.y = 1,
                 # hjust = -0.5,
                 # vjust = 1.5,
                 # font.label = list(size = 14, color = "black", face = "bold", family = NULL),
                 # widths = 1,
                 # heights = 1)# heights = c(0.25, 0.75))
  annotate_figure(figure,
                  top = text_grob(casecountsx$name[1], color = "blue", face = "bold", size = 14),
                  bottom = text_grob("Data source: \ C3.ai Covid Data Lake", color = "blue",
                                     hjust = 1, x = 1, face = "italic", size = 10),
                  left = text_grob("Data in Linear and Logarithmic Scales", color = "blue", rot = 90),
                  # right = "I'm done, thanks :-)!",
                  fig.lab = "Figure 2", fig.lab.face = "bold"
                  )
  
}
```
Checkin all data and plot casescounts2

```{r}
check_data(casecounts2)
plot_cases(casecounts2)
```

plot only logartihmic charts
```{r}
plot_cases_log<-function (casecountsx) {
  #### total as of today
  plot1 <- casecountsx %>% 
    ggplot(aes(dates, data/1000, color = value_id))+
    geom_line() +
    scale_y_log10(labels = waiver(), breaks= c(10,100,1000,10000,100000,1000000,10000000))+
    facet_wrap(vars(name)) +
    
   #theme_dark()
    #scale_color_viridis(discrete = FALSE, option = "C")+
    labs(
      x = "Date",
      y = "Count (Thousands)",
      color = "Metric"
    )
  
  figure <-ggarrange(plot1, ncol = 1,nrow = 1)
                 # label.y = 1,
                 # hjust = -0.5,
                 # vjust = 1.5,
                 # font.label = list(size = 14, color = "black", face = "bold", family = NULL),
                 # widths = 1,
                 # heights = 1)# heights = c(0.25, 0.75))
  annotate_figure(figure,
                  top = text_grob("Logorithmic Plot", color = "blue", face = "bold", size = 14),
                  bottom = text_grob("Data source: \ C3.ai Covid Data Lake", color = "blue",
                                     hjust = 1, x = 1, face = "italic", size = 10),
                  left = text_grob("Data in Logarithmic Scales", color = "blue", rot = 90),
                  # right = "I'm done, thanks :-)!",
                  fig.lab = "Figure 2", fig.lab.face = "bold"
                  )
  
}
```


```{r}
#check_data(casecountList)
plot_cases_log(casecounts2)
```


Export case counts as a .csv file.

```{r}
# Uncomment the line below to export the Tibble as a .csv file
# casecounts %>% write_csv("casecounts.csv")
```


### Mobility

Mobility data from Apple and Google provide a view of the impact of COVID-19 and social distancing on mobility trends.
```{r}
mobility_trends <- evalmetrics(
  "outbreaklocation",
  list(
    spec = list(
      ids = list("Miami-Dade_Florida_UnitedStates"),
      expressions = list(
        "Apple_WalkingMobility", 
        "Apple_DrivingMobility",
        "Google_ParksMobility",
        "Google_ResidentialMobility"
      ),
      start = "2020-03-01",
      end = "2020-07-31",
      interval = "DAY"
    )
  ),
  get_all = TRUE
)

mobility_trends
```

Plot these mobility trends.
```{r}
mobility_trends %>% 
  ggplot(aes(dates, data/100, color = value_id)) +
  geom_hline(aes(yintercept = 1), linetype = "dashed") +
  geom_line() +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(vars(name)) +
  labs(
    x = "Date",
    y = "Mobility compared to baseline",
    color = "Metric"
  )
```

### Projections

Use the GetProjectionHistory API to retrieve versioned time series projections for specific metrics made at specific points in time.
```{r}
# Retrieve projections made between April 13 and May 1 of mean total cumulative deaths in Miami Dade from April 13 to May 13
projections <- getprojectionhistory(
  list(
    outbreakLocation = "UnitedStates",
    metric = "UniversityOfWashington_TotdeaMean_Hist",
    metricStart = "2020-06-13", 
    metricEnd = "2020-07-31",
    observationPeriodStart = "2020-06-13",
    observationPeriodEnd = "2020-07-31"
  )
)

head(projections)
```

```{r}
# Retrieve actual total cumulative deaths in MiamiDade  from April 1 to July 31st
Miami_deaths <- evalmetrics(
  "outbreaklocation",
  list(
    spec = list(
      ids = list("Miami-Dade_Florida_UnitedStates"),
      expressions = list("JHU_ConfirmedDeaths"),
      start = "2020-04-01",
      end = "2020-07-31",
      interval = "DAY"
    )
  )
)

Miami_deaths

Miami_cases <- evalmetrics(
  "outbreaklocation",
  list(
    spec = list(
      ids = list("Miami-Dade_Florida_UnitedStates"),
      expressions = list("JHU_ConfirmedCases"),
      start = "2020-04-01",
      end = "2020-07-31",
      interval = "DAY"
    )
  )
)

US_deaths <- evalmetrics(
  "outbreaklocation",
  list(
    spec = list(
      ids = list("UnitedStates"),
      expressions = list("JHU_ConfirmedDeaths"),
      start = "2020-04-01",
      end = "2020-07-31",
      interval = "DAY"
    )
  )
)
Miami_cases
cor(Miami_cases$data,Miami_deaths$data)
```

Plot the results.
```{r}

ggplot() +
  geom_line(
    aes(dates, data), 
    data = Miami_deaths, 
    color = "red",

  ) +
  # geom_smooth(
  #   data=Miami_deaths,
  #   aes(dates, data),
  #   fill="blue",
  #   colour="darkblue", size=1
  #   )+ 
# 
  # geom_line(
  #   aes(dates, data),
  #   data = US_deaths,
  #   color = "black",
  # 
  # ) +
  facet_wrap(vars(name)) +
  labs(
    x = "Date",
    y = "Count",
    color = "Metric",
    title = "Cumulative death count"
    
  )
```


```{r}


